Meta-analytic findings of the self-controlled motor learning literature: Underpowered, biased, and lacking evidential value

### 2021-11-27 docs/accepted/manuscript-accepted.Rmd
update: added calculation for total N; set seed for z-curve; updated manuscript

abstract: |
  The self-controlled motor learning literature consists of experiments that compare a group of learners who are provided with a choice over an aspect of their practice environment to a group who are yoked to those choices. A qualitative review of the literature suggests an unambiguous benefit from self-controlled practice. A meta-analysis was conducted on the effects of self-controlled practice on retention test performance measures with a focus on assessing and potentially correcting for selection bias in the literature, such as publication bias and *p*-hacking. First, a naïve random effects model was fit to the data and a moderate benefit of self-controlled practice, *g* = .44 (*k* = 52, *N* = 3134, 95% *CI* [.31, .56]), was found. Second, publication status was added to the model as a potential moderator, revealing a significant difference between published and unpublished findings, with only the former reporting a benefit of self-controlled practice. Third, to investigate and adjust for the impact of selectively reporting statistically significant results, a weight-function model was fit to the data with a one-tailed *p*-value cutpoint of .025. The weight-function model revealed substantial selection bias and estimated the true average effect of self-controlled practice as *g* = .107 (95% *CI* [.047, .18]). *P*-curve analyses were conducted on the statistically significant results published in the literature and the outcome suggested a lack of evidential value. Fourth, a suite of sensitivity analyses were conducted to evaluate the robustness of these results, all of which converged on trivially small effect estimates. Overall, our results suggest the benefit of self-controlled practice on motor learning is small and not currently distinguishable from zero.

@@

## Naïve random effects model
The naïve random effects model estimated the average treatment effect of self-controlled practice, *g* = .44 (*k* = 52, *N* = 3134, 95% *CI* [.31, .56]).  However, there was significant variability in the average effect estimated across experiments, *Q*(*df* = 51) = 103.45, *p* < .0001, $\tau$ = .31. It was estimated that 47.9% ($I^2$) of the total variability in effect sizes across experiments was due to true heterogeneity in the underlying effects measured (see Figure \@ref(fig:fig3)).

@@

## *Z*-curve
A *z*-curve was fit to the overall retention data and estimated the power of statistically significant studies (ERR) as 14%, 95% CI [5%, 35%]. The power of all studies conducted (EDR) was estimated as 6%, 95% *CI* [5%, 14%]. The 95% confidence intervals for both the ERR and EDR failed to include the observed discovery rate of 48%, suggesting significant publication bias in the data.

>>>>

abstract: |
  The self-controlled motor learning literature consists of experiments that compare a group of learners who are provided with a choice over an aspect of their practice environment to a group who are yoked to those choices. A qualitative review of the literature suggests an unambiguous benefit from self-controlled practice. A meta-analysis was conducted on the effects of self-controlled practice on retention test performance measures with a focus on assessing and potentially correcting for selection bias in the literature, such as publication bias and *p*-hacking. First, a naïve random effects model was fit to the data and a moderate benefit of self-controlled practice, *g* = .44 (*k* = 52, *N* = 2061, 95% *CI* [.31, .56]), was found. Second, publication status was added to the model as a potential moderator, revealing a significant difference between published and unpublished findings, with only the former reporting a benefit of self-controlled practice. Third, to investigate and adjust for the impact of selectively reporting statistically significant results, a weight-function model was fit to the data with a one-tailed *p*-value cutpoint of .025. The weight-function model revealed substantial selection bias and estimated the true average effect of self-controlled practice as *g* = .107 (95% *CI* [.047, .18]). *P*-curve analyses were conducted on the statistically significant results published in the literature and the outcome suggested a lack of evidential value. Fourth, a suite of sensitivity analyses were conducted to evaluate the robustness of these results, all of which converged on trivially small effect estimates. Overall, our results suggest the benefit of self-controlled practice on motor learning is small and not currently distinguishable from zero.

@@

## Naïve random effects model
The naïve random effects model estimated the average treatment effect of self-controlled practice, *g* = .44 (*k* = 52, *N* = 2061, 95% *CI* [.31, .56]).  However, there was significant variability in the average effect estimated across experiments, *Q*(*df* = 51) = 103.45, *p* < .0001, $\tau$ = .31. It was estimated that 47.9% ($I^2$) of the total variability in effect sizes across experiments was due to true heterogeneity in the underlying effects measured (see Figure \@ref(fig:fig3)).

@@

## *Z*-curve
A *z*-curve was fit to the overall retention data and estimated the power of statistically significant studies (ERR) as 12%, 95% CI [3%, 34%]. The power of all studies conducted (EDR) was estimated as 6%, 95% *CI* [5%, 13%]. The 95% confidence intervals for both the ERR and EDR failed to include the observed discovery rate of 48%, suggesting significant publication bias in the data.
